A Handbook for Tenured Track Faculty in Computer Science in the US

### 2024-07-08 faculty.tex
updates

Many would advise about being strategic about your services.I think it is a good advice as you can get more out of your services. However, I do not follow this, and in many cases I would volunteer for tasks that no one wants to do because I feel that someone gotta do it and I can help out.

@@

Toward the end of your 3rd year, you will need to prepare a renewal package and submit it to your department.  The department will then review your package and vote on whether to renew you for your 4th (and perhaps also 5th year). Your next renewal is essentially your tenure application. The department will vote on whether to recommend your renewal.

Similarly, after 5 or 6 years, it is time for you to go up (or apply) for tenure.  You will need to prepare a tenure package and submit it to your department.  The department will then review your package and vote on whether to recommend you for tenure.  If the department votes in your favor, the package will then go to the college, and then to the university (e.g., provost, president, board of visitors, etc).  The university will then make the final decision on whether to grant you tenure.

@@

\section{Timeline}
Typically you will first mention to your department chair that you are planning to go up for tenure. The chair might then assign you a mentor who will help you with the process. You will apply for tenure in the Fall semester, the decision will be made in the next Spring semester, and you will get tenure in the Fall semester after that.

@@

\begin{itemize}
  \item Fall 2021: Moved to GMU with accelerated tenure clock (meaning I don't have to restart from the beginning as I have already been at UNL since 2016).
  \item (late) Fall 2022: I told my department chair that I plan to apply for tenure in 2023 (which means if things go well then I get tenure in Fall'24).
  \item Spring 2023: I started preparing my tenure package, which includes writing statements and think reference letters.  I also took a semester off because GMU has a pre-tenure study break policy.
  \item Summer 2023: I submitted my tenure package.
  \item Fall 2024: I gave a tenure talk. Department voted in my favor, and then the college voted in my favor, and the decision goes up to the university.
  \item Spring 2024: In May, I got the official notification that I got tenure and promoted to Associate Prof., beginning in Fall 2024.
\end{itemize}

@@

GMU (and likely other universities) has a very strict and specific format for the CV (e.g., what to include, how to format, etc).  Fortunately I have been using this format since I joined GMU and also kept my CV up-to-date, so I didn't have to do much. I did have to add a few more sections (e.g., teaching evaluation and number of students in courses taught).  I also have to make sure that the CV is consistent with my statements and other parts of the package (e.g., funding amount must match with what the university has in their system).

I also looked at CVs of other people at GMU who recently got tenure. I also sent my CV to  my mentors and got some feedback.

@@

Here is my \href{https://dynaroars.github.io/people/nguyenthanhvuh/files/pt/statement-annotated.pdf}{tenure application statement}, annotated with comments. At GMU, the statement is limited to 8 pages.  My statement follows the standard structure and talks about my achievements and plans in (i) research and funding, (ii) advising and mentoring, (iii) teaching, and (iv) services. The main unique part of my statement is the first page where I highlight and summarize my achievements (e.g., papers, fundings, students, etc). I get most of these information and numbers from my CV.

While preparing the statement was the most time-consuming part of my tenure package, it didn't take as long (may be just a couple of days).  I get statements from several colleague at GMU who recently got tenure and a couple of people at other universities to get some ideas on how it should be structured.  I was able to reuse lots of text and contents from my job application statements (when I applied to GMU in 2021).  I did not send it for feedback and also did not revise it much.

>>>>

Many would (rightly) advise about being strategic about your services. I do not really generally this and just volunteer for services that I think it is my responsibility (e.g., being in PCs, though I never liked them much) and those that I enjoy (NSF panels, services to the depts). In many cases I would also volunteer for tasks that no one wants to do because I feel that someone gotta do it and I can help out, e.g., various misc tasks in the dept.

@@

Toward the end of your 3rd year (or early 4th year), you will need to prepare and submit a renewal package.  The department will then review your package and vote on whether to renew you for your 4th (and perhaps also 5th year). Many faculty members actually do not get renewed, so do take this process seriously.

Then, after 5 or 6 years, it is time for you to go up (or apply) for tenure.  You will need to submit a tenure package.  The department will then review your package and vote on whether to recommend you for tenure.  If the department votes in your favor, the package will then go to the college, and then to the university (e.g., provost, president, board of visitors, etc).  The university will then make the final decision on whether to grant you tenure (typically they follow the department's recommendation, but not always).

I will focus on the tenure application process, as the renewal process is similar but less intense (e.g., no external letters).

@@

\section{Timeline}

If you plan to apply for tenure in year X (and get tenure in year X+1), you will first mention your intention to the dept chair (at GMU this is at least a semester before, i.e., in December of year X-1). The chair might then assign you a mentor who will help you with the process. You will apply for tenure in the summer or early Fall of year X, the decision will be made in the next Spring semester of year X+1, and if everything goes well, you will get tenure in the Fall semester in year X+1.

@@

\begin{itemize}
  \item Fall 2021: Moved to GMU with accelerated tenure clock (meaning I don't have to restart from the beginning as I have already been at UNL since 2016).
  \item (late) Fall 2022: Told my dept. chair that I plan to apply for tenure in 2023.
  \item Spring 2023: Prepared my tenure package, which includes writing statements and think reference letters.  Also took this semester off (no teaching no services) because GMU has a pre-tenure study break policy.
  \item Summer 2023: Submitted my tenure package.
  \item Fall 2024: Gave a tenure talk. Department voted in my favor, and then the college voted in favor, and the decision goes up to the university.
  \item Spring 2024: In May, got the official notification that I got tenure and promoted to Associate Prof., beginning in Fall 2024.
\end{itemize}

@@

Both GMU and UNL ( and likely other universities) have a very strict and specific format for the CV, e.g., what to include, how to format, order of sections and subsections, etc.  
Fortunately I have always been using the university CV format and also kept my CV up-to-date, so I didn't have to do much. 
I did have to add more \emph{personal} details (e.g., teaching evaluation and funding amounts that match what the university has in their system).

I also looked at CVs of other people in the dept. who recently got tenure. I also sent my CV to  my mentors and got many feedback. In fact, I also asked a Division Dean in the College who is in charge of P\&T procedure to review my CV for formatting and other issues. In short, the CV is one of the easiest part of the tenure package, and so I want to make sure mine has no silly mistakes.

@@

Here is my \href{https://dynaroars.github.io/people/nguyenthanhvuh/files/pt/statement-annotated.pdf}{tenure application statement}, annotated with comments. At GMU, the statement is limited to 8 pages.  My statement talks about my achievements and plans in (i) research and funding, (ii) advising and mentoring, (iii) teaching, and (iv) services. This organization is somewhat standard, and the main unique part of my statement is the first page where I highlight and summarize my achievements (e.g., papers, fundings, students, etc). I got these information and numbers from my CV.

While preparing the statement was the most time-consuming part of my tenure package, it wasn't too bad and may be just took a couple of days.  I was able to reuse lots of text and contents from my job application statements (when I applied to GMU in 2021). I also got statements from colleagues at GMU who recently got tenure and a couple of people at other universities to get some ideas on what to write.  I did not send my statement for feedback because I felt that it was good enough and I didn't want to bother people (reading an 8-page statement is a lot more work).
